---
title: "Initial Notes"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Rgex Libs

```{r}
library(stringr)
```

# Examples

```{r}
animals <- c("jaguar", "jegue", "gato")

str_detect(animals, pattern = "t")

str_extract(animals, pattern = "t")

str_locate(animals, pattern = "to")

```

There are 14 meta chars that carry special meaning inside regular expressions. So we need to escape them with a backslash if we want to match the literal char.

```{r}
math_chars <- c("1 + 1", "2 + 5", "2 -2")
str_detect(math_chars, "\\+")
str_extract(math_chars, "\\-")

```


# Wildcards

## The Dot

This means that this character will match anything in place other then a newline character.

```{r}
strings <- c("cat", "dog", "cut")

str_extract(strings, "c.")


```


## Character classes

A character class is constructed with square brackets and allows us to match a character specified inside the class.

```{r}
str_extract(strings, "[at]")

```

White spaces inside character classes are meaningful as they are interpreted as literal chars.

```{r}
string_w_space <- c("cat ", "c at")

str_extract_all(string_w_space, "[a t]")


```


We can use a hyphen to set a range of chars.

```{r}
strings <- c("1", "3", "4", "6")

str_extract(strings, "[1-5]")

```


```{r}
sentence <- "This is a long sentence with a number 3 on in it"

str_extract_all(sentence, "[1-3a-b]")
str_locate_all(sentence, "[1-3a-b]")

```

We can also negate characters in a class with a caret ^.

```{r}
strings_more <- c("a", "b", "c")

str_extract(strings_more, "[^ac]")
```

This will match anything that isn't the letter a or c.


## Shorthand Char

Certain character classes are so commonly used that they have been predefined with names. 

- [:digit:]: all the digits
- \\d: the same
- [:alpha:]: alphabetic characters, upper and lower
- [:alnum:]
- [:graph:]: graphical chars, the same of [[:alnum:][:punct:]]
- [:space:]: space, tab, newline, form, feed, carriage, \\s
- [^[:space:]] or \\S: not a space


```{r}
string2 <- c("frase with a spa c e")

str_extract_all(string2, pattern = "[^[:space:]]")
str_extract_all(string2, pattern = "\\S")


```


## Quantifiers

Specify how many times we expect something to occur using quantifiers.

Notice that:

- ? is a shorthand of {0, 1} => zero or one time
- + is a shorthand of {1, } => one or more times
- * is a shorthand of {0, } => zero or more times
- {n} => exactly n times
- {n, } => at least n times
- {n, m} => between n and m times


```{r}
string_quant <- c("color", "colour", "colours")


str_detect(string_quant, "colou?r[s]?")
str_extract(string_quant, "colou?r[s]?")

# extract a sequence of numbers
string_quant_2 <- "This sentence we have some data year 1998."

str_extract(string_quant_2, pattern = "\\d{4}")
str_extract(string_quant_2, pattern = "[:digit:]{4}")

```


Sometimes we want the repetition to happen over multiple characters.


```{r}
# we want to stay with NA BATMAN!!!
batman <- "NA NA NA NA NA NA NA NA NA NA NA NA NA NA BATMAN!!!"

str_remove(batman, "(NA )+")
str_remove(batman, "(NA )*(NA){1}")

```

By default these matches are greedy (try to match the longest string possible). We can instead make them lazy by placing a ? after. 

- ??: zero or one times, prefers 0
- *?: zero or more times, match as few times as possible
- +?: one or more times, match as few times as possible
- {n}? exactly n times, match as few times as possible
- {n,}? at least n times, match as few times as possible
- {n, m}? between n and m times, match as few times as possible but at least n

```{r}
str_extract(batman, "(NA ){3,7}")
str_extract(batman, "(NA ){3,7}?")

```


## Anchors

- ^: first element
- $: last element

_First you get the power ^ and then you get the money $_


```{r}
seasons <- c("The summer is hot this year",
             "The spring is a lovely time",
             "Winter is my favorite time of the year",
             "Fall is a time of peace")
str_detect(seasons, "^The")
str_extract(seasons, "^The")
str_extract(seasons, "year$")


```

Combining the two anchors we can get the specific word.

```{r}
string_achors <- c("data", "data-raw", "data science")

str_extract(string_achors, "^data$")

```





















































