---
title: "Initial Notes"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Rgex Libs

```{r}
library(stringr)
```

# Examples

```{r}
animals <- c("jaguar", "jegue", "gato")

str_detect(animals, pattern = "t")
str_extract(animals, pattern = "t")
str_locate(animals, pattern = "to")

```

There are 14 meta chars that carry special meaning inside regular expressions. So we need to escape them with a backslash if we want to match the literal char.

```{r}
math_chars <- c("1 + 1", "2 + 5", "2 -2")
str_detect(math_chars, "\\+")
str_extract(math_chars, "\\-")

```


# Wildcards

## The Dot

This means that this character will match anything in place other then a newline character.

```{r}
strings <- c("cat", "dog", "cut")

str_extract(strings, "c.")


```


## Character classes

A character class is constructed with square brackets and allows us to match a character specified inside the class.

```{r}
str_extract(strings, "[at]")

```

White spaces inside character classes are meaningful as they are interpreted as literal chars.

```{r}
string_w_space <- c("cat ", "c at")

str_extract_all(string_w_space, "[a t]")


```


We can use a hyphen to set a range of chars.

```{r}
strings <- c("1", "3", "4", "6")

str_extract(strings, "[1-5]")

```


```{r}
sentence <- "This is a long sentence with a number 3 on in it"

str_extract_all(sentence, "[1-2a-b]")
str_locate_all(sentence, "[1-2a-b]")

```

We can also negate characters in a class with a caret ^.

```{r}
strings_more <- c("a", "b", "c")

str_extract(strings_more, "[^ac]")
```

This will match anything that isn't the letter a or c.


## Shorthand Char

Certain character classes are 